drivers response diversity

### 2024-09-26 03_drivers_response_diversity_Fig3_Fig4.Rmd
renamed script
preparing the code for figure 4 a and c

# select the columns we use
lake_selection <- lake_info |> 
  select(Lake, Phos_max, Max_depth, Lake_area)

@@

# exclude lepomis gibbosus and re-calculate species richness per lake
species_endemism_short <- species_endemism |> 
  filter(species != "Lepomis_gibbosus") |> 
  select(-sum_species) |> 
  group_by(fLake) |> 
  mutate(sum_species = sum(num_species))

@@

# phosphorus values as numeric
df_lm$Phos_max <- as.numeric(df_lm$Phos_max)

@@

# save as TIFF
# tiff(paste("total_models/plots/lm_divergence_all.tiff", sep = ""), units="in", width=11, height=7, res=300)
# # plot(ggarrange(depth1, depth2, ncol = 2))
# # plot science discussion
# plot(lm_divergence_all)
# 
# dev.off()
```

## Figure 4
```{r}
# second lm for maximum phosphorus ~ species richness
lm2 <- lm(mean_rdiv ~ Phos_max + sum_species, data = df_lm_excl)

@@

@@

```
### OLD CODE Community subgroups
####Fig. 4a
We fitted simple linear models for the number of species in each category and the mean dissimilarity in each lake. 
```{r}
# Lms for all categories: endemic, translcated, non_native and non_endemic_native
# with data on all lakes
# Endemic
lm_endemic <- lm(mean_rdiv ~ endemic, data = df_lm)

>>>>

# select the columns we use
lake_selection <- lake_info |> 
  select(Lake, Phos_max, Max_depth, Lake_area, Species_loss, Phos_max_Vonlanthen)

@@

# re-calculate species richness per lake
species_endemism_short <- species_endemism |> 
  # filter(species != "Lepomis_gibbosus") |> 
  select(-sum_species) |> 
  group_by(fLake) |> 
  mutate(sum_species = sum(num_species))

@@

# phosphorus values as numeric
df_lm$Phos_max <- as.numeric(df_lm$Phos_max)
df_lm$Phos_max_Vonlanthen <- as.numeric(df_lm$Phos_max_Vonlanthen)

@@

# save as TIFF
# tiff(paste("total_models/plots/lm_divergence_all.tiff", sep = ""), units="in", width=11, height=7, res=300)
# # plot(ggarrange(depth1, depth2, ncol = 2))
# # plot science discussion
# plot(lm_divergence_all)
# 
# dev.off()
```
## Figure 4a
```{r}
df_lm_4 <- df_lm |> 
  filter(!Lake %in% c("Poschiavo", "Joux", "Maggiore", "Lugano"))

lm_species_loss <- lm(Species_loss ~ Phos_max_Vonlanthen, data = df_lm_4)

# checking model assumptions
summary(lm_species_loss)
shapiro.test(resid(lm_species_loss))
lmtest::bptest(lm_species_loss)
plot(lm_species_loss)

# new data
new_data_species_loss <- tibble(Phos_max_Vonlanthen = seq(from = min(df_lm_4$Phos_max_Vonlanthen), to = max(df_lm_4$Vonlanthen), length = 15))

# predict model
prediction_species_loss <- predict.lm(lm_species_loss, newdata = new_data_species_loss, se.fit = TRUE, type = "response")

# combine new data and predictions
bind_model_species_loss <- cbind(new_data_species_loss, prediction_species_loss) |> 
  mutate(category ="color")

bind_model_species_loss|> 
  ggplot(aes(Phos_max, fit, color = category, fill = category)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = (fit - se.fit), ymax = (fit + se.fit)),  alpha = 0.3, color = NA) +
  theme_classic(base_size = 20) +
  ylab("Species loss") +
  xlab("Maximum total phosphorus (µg/L)") +
  # ylim(1,3) +
  scale_fill_manual(values = mycolor, aesthetics = c("color", "fill"), guide = NULL)

```
## Figure 4b
```{r}
# second lm for maximum phosphorus ~ species richness
lm2 <- lm(mean_rdiv ~ Phos_max + sum_species, data = df_lm_excl)

@@

##Figure 4c
```{r}

lm_test <- lm(mean_rdiv ~ Species_loss, data = df_lm_4)

# checking model assumptions
summary(lm_test)
# shapiro.test(resid(lm_species_loss))
# lmtest::bptest(lm_species_loss)
# plot(lm_species_loss)

# new data
new_data_test <- tibble(Species_loss = seq(from = min(df_lm_4$Species_loss), to = max(df_lm_4$Species_loss), length = 15))

# predict model
prediction_test <- predict.lm(lm_test, newdata = new_data_test, se.fit = TRUE, type = "response")

# combine new data and predictions
bind_model_test <- cbind(new_data_test, prediction_test) |> 
  mutate(category ="color")

bind_model_test |> 
  ggplot(aes(Species_loss, fit, color = category, fill = category)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = (fit - se.fit), ymax = (fit + se.fit)),  alpha = 0.3, color = NA) +
  theme_classic(base_size = 20) +
  ylab("Mean dissimilarity") +
  xlab("Species loss") +
  # ylim(1,3) +
  scale_fill_manual(values = mycolor, aesthetics = c("color", "fill"), guide = NULL)

@@

```
# OLD CODE Community subgroups
####Fig. 4a
We fitted simple linear models for the number of species in each category and the mean dissimilarity in each lake. 
```{r}
# Lms for all categories: endemic, translcated, non_native and non_endemic_native
# with data on all lakes
# Endemic
lm_endemic <- lm(mean_rdiv ~ endemic, data = df_lm)
